---
title: 'Descriptive Statistics'
author: "Chunyi Wu and Andrzej Galecki"
date: "`r as.character(Sys.Date(), format = '%A %B %d, %Y')`"
output:
  rmdformats::readthedown:
     lightbox: true
     use_bookdown: true
---


```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```


```{r pacman, echo=FALSE, message=FALSE, warning=FALSE}
if (!require('pacman')) install.packages('pacman', repos = "http://cran.us.r-project.org")
library(pacman)

pacman::p_load(
  readxl,       # load excel file
  glmnet,       # Lasso and Elastic-Net Regularized Generalized Linear Models
  rmdformats,   # rmd formats
  rmarkdown,    # rmarkdown
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics 
  janitor,      # adding totals and percents to tables
  flextable,    # format the output table
  gtsummary,    # calculate summary statistics and format the results
  sjPlot,       # correlation matrix
  purrr,        # enhances R functional programming (FP) toolkit 
  tidyr,        #Tools to help to create tidy data
  ggplot2,      #Plot the results
  corrplot
  )
```

# R setup

```{r  rsetup}
sessionInfo()
```

# Univariate statistics


```{r, data-read, echo=FALSE, message=FALSE, warning=FALSE}
xlsx_path = "./Data/data_example.xlsx"

scrambled_orig=readxl::read_excel(xlsx_path, na = "", guess_max = 1000)
scrambled <- scrambled_orig  %>% mutate(log10_B_HBA1C_PRC=log10(B_HBA1C_PRC),log10_DU_ACR=log10(DU_ACR))
```
Raw data stored in `r xlsx_path` file contain `r nrow(scrambled_orig)` rows.  
Variable names are listed below.

```{r var-names}
names(scrambled_orig)
```

```{r, table1, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
scrambled2 <- scrambled  %>% 
          mutate(Gender=case_when(SEX==1 ~"Male",SEX==2 ~"Female"),
              Status=case_when(CASE_CONTROL==1 ~"Case",CASE_CONTROL==0 ~"Control"))%>%                                
          select(-INDEX,-SEX,-CASE_CONTROL) %>%
          select("Gender","Status","Death" , "FU_TIME" ,"FU_TIME_Death","SBP_TL","DBP_TL" , 
                 "AGEONSET_TL" , "BMI_TL","AGE_TL" , "B_HBA1C_PRC","log10_B_HBA1C_PRC", 
                 "DU_ACR","log10_DU_ACR","BL_eGFR","KIM1.npx" ,"SYND1.npx","IL.1RT1.npx",
                 "WFDC2.npx",      "CD27.npx" , "TNFRSF10A.npx","LAYN.npx","PVRL4.npx", 
                 "EDA2R.npx" ,"TNFRSF4.npx" ,"GFR_alpha_1_npx", "TNF.R1.npx", "PI3_npx" , 
                 "EFNA4.npx","TNF.R2.npx" ,"DLL1.npx" ,"TNFRSF6B.npx" , 
                 "CD160.npx","EPHA2.npx","RELT.npx","LTBR.npx")         

scrambled2$FU_TIME_Death=as.numeric(scrambled2$FU_TIME_Death)

scrambled2 %>%
  tbl_summary(by = Status, missing_text = "(Missing N)",   label = list(FU_TIME ~ "Follow-up Time (year)", FU_TIME_Death ~ "Follow-up Time until Death (year)", SBP_TL ~ "Baseline Systolic blood pressure (mm-Hg)", DBP_TL ~ "Baseline Diatolic blood pressure (mm-Hg)",AGEONSET_TL ~ "Age at T1D onset (year)", BMI_TL ~ "Baseline BMI (kg/m2)", AGE_TL ~ "Baseline Age (year)",B_HBA1C_PRC ~ "Baseline HbA1c (%)",DU_ACR ~ "Baseline ACR (mg/g Cr)",BL_eGFR ~"Baseline eGFR (mL/min/1/.73m2)",log10_B_HBA1C_PRC ~"log10 HbA1c", log10_DU_ACR ~"log10 ACR" ), type = all_continuous() ~ "continuous2", statistic = list(all_continuous() ~ c("{mean}, {sd}","{median} ({p25}, {p75})", "{min}, {max}"),all_categorical() ~ "{n} ({p}%)"))  %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>%
  add_overall() %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Status**") %>%
  modify_footnote( all_stat_cols() ~ "Mean (SD), Median (IQR) or Frequency (%)")%>%
  modify_caption("**Table 1. Patient Characteristics**") %>%
  bold_labels()
```


# Histograms

```{r, histogram, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 12, fig.height=8, dev ="svg"}
var_num <- scrambled  %>% select("FU_TIME" ,"FU_TIME_Death","SBP_TL","DBP_TL" , "AGEONSET_TL" , "BMI_TL","AGE_TL" , "B_HBA1C_PRC","log10_B_HBA1C_PRC", "DU_ACR","log10_DU_ACR","BL_eGFR","KIM1.npx" ,"SYND1.npx","IL.1RT1.npx","WFDC2.npx",      "CD27.npx" , "TNFRSF10A.npx","LAYN.npx","PVRL4.npx", "EDA2R.npx" ,"TNFRSF4.npx" ,"GFR_alpha_1_npx", "TNF.R1.npx", "PI3_npx" , "EFNA4.npx","TNF.R2.npx" ,"DLL1.npx" ,"TNFRSF6B.npx" , "CD160.npx","EPHA2.npx","RELT.npx","LTBR.npx") 

var_num %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(bins=20)

```

## Case

```{r, histogram_case, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 12, fig.height=8, dev ="svg"}
Case <- scrambled %>% filter (CASE_CONTROL==1) %>% select("FU_TIME" ,"FU_TIME_Death","SBP_TL","DBP_TL" , "AGEONSET_TL" , "BMI_TL","AGE_TL" , "B_HBA1C_PRC","log10_B_HBA1C_PRC", "DU_ACR","log10_DU_ACR","BL_eGFR","KIM1.npx" ,"SYND1.npx","IL.1RT1.npx","WFDC2.npx",      "CD27.npx" , "TNFRSF10A.npx","LAYN.npx","PVRL4.npx", "EDA2R.npx" ,"TNFRSF4.npx" ,"GFR_alpha_1_npx", "TNF.R1.npx", "PI3_npx" , "EFNA4.npx","TNF.R2.npx" ,"DLL1.npx" ,"TNFRSF6B.npx" , "CD160.npx","EPHA2.npx","RELT.npx","LTBR.npx") 

Case %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(bins=20)

```

## Control

```{r, histogram_control, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 12, fig.height=8, dev ="svg"}
Control <- scrambled %>% filter (CASE_CONTROL==0) %>% select("FU_TIME" ,"FU_TIME_Death","SBP_TL","DBP_TL" , "AGEONSET_TL" , "BMI_TL","AGE_TL" , "B_HBA1C_PRC","log10_B_HBA1C_PRC", "DU_ACR","log10_DU_ACR","BL_eGFR","KIM1.npx" ,"SYND1.npx","IL.1RT1.npx","WFDC2.npx",      "CD27.npx" , "TNFRSF10A.npx","LAYN.npx","PVRL4.npx", "EDA2R.npx" ,"TNFRSF4.npx" ,"GFR_alpha_1_npx", "TNF.R1.npx", "PI3_npx" , "EFNA4.npx","TNF.R2.npx" ,"DLL1.npx" ,"TNFRSF6B.npx" , "CD160.npx","EPHA2.npx","RELT.npx","LTBR.npx") 

Control %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(bins=20)

```

# Protein Results

## Spearman's Correlation Matrix

```{r proteins-vec, include=FALSE}
proteins <- c("KIM1.npx",      "SYND1.npx", "IL.1RT1.npx","WFDC2.npx", "CD27.npx",
              "TNFRSF10A.npx", "LAYN.npx",  "PVRL4.npx",  "EDA2R.npx" ,"TNFRSF4.npx" ,
              "GFR_alpha_1_npx", "TNF.R1.npx", "PI3_npx" , "EFNA4.npx","TNF.R2.npx" ,
              "DLL1.npx" ,"TNFRSF6B.npx" , "CD160.npx","EPHA2.npx","RELT.npx",
              "LTBR.npx") 
```

Spearman-correlation coefficients are estimated for the following proteins:

```{r proteins-list, echo=FALSE}
proteins
```

```{r Correlation-table-proteins, echo=FALSE, message=FALSE, warning=FALSE}
protein <- scrambled  %>% select(all_of(proteins))
tab_corr(protein, p.numeric=TRUE, corr.method = c( "spearman"))
```

### Case
```{r Correlation-table-proteins_c, echo=FALSE, message=FALSE, warning=FALSE}
protein1 <- Case  %>% select(all_of(proteins))
tab_corr(protein1, p.numeric=TRUE, corr.method = c( "spearman"))
```

### Control
```{r Correlation-table-proteins_n, echo=FALSE, message=FALSE, warning=FALSE}
protein2 <- Control  %>% select(all_of(proteins))
tab_corr(protein2, p.numeric=TRUE, corr.method = c( "spearman"))
```

## Cluster dendrogram

```{r Correlation-proteins, echo=FALSE, message=FALSE, warning=FALSE,fig.width = 12, fig.height=8, dev ="svg"}

cors <- cor(protein, method = c( "spearman"), use ="complete.obs")
hcl <- hclust(as.dist(1-abs(cors)))
plot(hcl, xlab ="Proteins", sub ="")
```

From `hclust` documentation: "In hierarchical cluster displays, a decision is needed at each merge to specify 
which subtree should go on the left and which on the right. Since, for \(n\) observations there are \(n-1\) merges,
there are \(2^{(n-1)}\) possible orderings for the leaves in a cluster tree, or dendrogram. The algorithm used 
in hclust is to order the subtree so that the tighter cluster is on the left (the last, i.e., most recent, 
merge of the left subtree is at a lower value than the last merge of the right subtree). Single observations
are the tightest clusters possible, and merges involving two observations place them in order by their 
observation sequence number."

### Case

```{r Correlation-proteins_c, echo=FALSE, message=FALSE, warning=FALSE,fig.width = 12, fig.height=8, dev ="svg"}

cors1 <- cor(protein1, method = c( "spearman"), use ="complete.obs")
hcl1 <- hclust(as.dist(1-abs(cors1)))
plot(hcl1, xlab ="Proteins", sub ="")
```

### Control

```{r Correlation-proteins_n, echo=FALSE, message=FALSE, warning=FALSE,fig.width = 12, fig.height=8, dev ="svg"}

cors2 <- cor(protein2, method = c( "spearman"), use ="complete.obs")
hcl2 <- hclust(as.dist(1-abs(cors2)))
plot(hcl2, xlab ="Proteins", sub ="")
```

## Visualization of the correlation matrix

```{r Corr-proteins-ord, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 12, fig.height=8, dev ="svg",out.extra='style="max-width:none; width:100vw; margin-left:calc(50% - 50vw);"'}

ord <- hcl$order
cors_ord <- cors[ord,ord]
cors_ord <- round(cors_ord,2)
corrplot.mixed(cors_ord, upper ="pie", lower="number", tl.pos ="lt",
 mar = c(0,0,1,0), tl.cex=0.7, main="Correlation matrix plot", number.cex = 0.8,
 lower.col="black")
```

`out.extra` option based on (https://stackoverflow.com/questions/67630290/allow-plots-to-overhang-chunk-text-in-rmarkdown-html)

### Case

```{r Corr-proteins-ord_c, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 12, fig.height=8, dev ="svg",out.extra='style="max-width:none; width:100vw; margin-left:calc(50% - 50vw);"'}

cors_ord1 <- cors1[ord,ord]
cors_ord1 <- round(cors_ord1,2)
corrplot.mixed(cors_ord1, upper ="pie", lower="number", tl.pos ="lt",
 mar = c(0,0,1,0), tl.cex=0.7, main="Correlation matrix plot", number.cex = 0.8,
 lower.col="black")
```

### Control

```{r Corr-proteins-ord_n, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 12, fig.height=8, dev ="svg",out.extra='style="max-width:none; width:100vw; margin-left:calc(50% - 50vw);"'}

cors_ord2 <- cors2[ord,ord]
cors_ord2 <- round(cors_ord2,2)
corrplot.mixed(cors_ord2, upper ="pie", lower="number", tl.pos ="lt",
 mar = c(0,0,1,0), tl.cex=0.7, main="Correlation matrix plot", number.cex = 0.8,
 lower.col="black")
```

```{r scrambled-RDS, echo=FALSE, message=FALSE, warning=FALSE}
pathRDS <- "./Data/scrambled.rds"
scrambled <- scrambled %>% mutate(status=CASE_CONTROL, time=FU_TIME)
saveRDS(scrambled, file = pathRDS)
```

Note: Data stored in object `scrambled` are saved in `r pathRDS` file.

