---
title: 'Multiple Cox Models'
author: "Chunyi Wu and Andrzej Galecki"
date: "`r format(Sys.Date())`"
output:
  rmdformats::readthedown:
    lightbox: true
    use_bookdown: true
params:
   ei: 4 
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
ei <- params$ei #
print(ei)

```

# LASSO (- `r ei` :=  `r xvars[ei]`)

Covariate # `r ei` :=  `r xvars[ei]` removed 


```{r datain-info}
excel_datain  # Path to Excel input raw dataset
prot_npx 
clin_vars
surv_vars
dim(x)
tibble(foldid = foldid) %>% group_by(foldid) %>% count()

```

<!-- glmnet object -->


## glmnet 


```{r glmnet-object}
ei
xvars[ei]
ncolx <- length(colnames(x))
pen <- rep(1, times= ncolx)
pen[1:length(clin_vars)] <- 0


fit <- glmnet(x, ySurv, family = "cox", alpha=1, penalty.factor = pen, exclude = ei)
(gl_fit <- myglance(fit))
(td_fit <- mytidy(fit))
plot(fit)
```

```{r  ggplot-coefs}
dfg <- td_fit %>% unnest(beta) %>% group_by(step) 

ggplot(dfg , aes(x = log(lambda), y = estimate)) +
       geom_line(aes(color = term)) + ylab("Coefficients")
```

## cv.glmnet 

```{r cv-glmnet}
cvfit <- cv.glmnet(x, ySurv, family = "cox", alpha=1, foldid = foldid, penalty.factor = pen, exclude =ei)
cvfit
# mytidy(cvfit)
(mygl <- myglance(cvfit))
lmbda <- mygl[,c("lambda.min", "lambda.1se")] %>% 
           mutate(log_lmin = log(lambda.min), log_l1se = log(lambda.1se))
lmbda

## cv.plot
#mytidy(cvfit)
plot(cvfit)
```

## c-v partial likelihood dev

* cross-validated  partial likelihood deviance



```{r td-cvfit1}
td_cvfit <- mytidy_Surv(cvfit, x, ySurv)
td_cvfit$comment <- paste0(xvars[ei], " removed") 
td_cv <-td_cvfit  %>% select(-c(beta, conf.low, conf.high, df, AICc)) 
td_cv %>% filter(mincv != "-") %>% print(n = 500)
```
## beta coefficients

```{r betai}
betai <- td_cvfit %>% select(c(step,lambda, nzero, beta, mincv, comment)) %>% 
           unnest(beta) %>% group_by(step)
nrow(betai)
betai %>% filter(mincv != "-") %>% print(n=400)
```


```{r coef-beta}
bx <- betai %>% select(term, estimate, mincv)
bx %>% filter(mincv == "1se" | mincv == "min" | mincv == "+")
```


## Coefficients path

* Plot using glmnet fit object

```{r beta-path}
plot(fit, xvar = "lambda", label = TRUE)
```


* Template for coefficients.
* Rows correspond to (descending) sequence of lambda hyperparameters
* Columns correspond to predictors

```{r tcoefs1}
coefs <- coef(fit)
ncoefs <- nrow(coefs)
snms <- colnames(coefs)
idx_1se <- mygl %>% select(index_1se) %>% pull()
idx_min <- mygl %>% select(index_min) %>% pull()

snms[idx_1se] <- paste0("<" ,snms[idx_1se])
snms[idx_min] <- paste0(snms[idx_min], ">")

tx <- t(coefs)
ttx <- t(tx)
tx0 <- t(ttx != 0)
idx <- apply(tx0, 1, function(x) paste(ifelse(x, 1, 0), collapse =""))
idxr <- rank(idx)
tmp <- idxr[c(idx_1se, idx_min)] 
idxr[c(idx_1se, idx_min)] <- tmp +0.01
sel <- which(!duplicated(idxr))
res <- tx0[sel,]
rownames(res) <- snms[sel]
```

```{r tcoefs2}
cnms <- 1:ncoefs
tmp <- paste0( ei, "*")
cnms[ei] <- tmp 
colnames(res) <- cnms
```
* Predictors 

```{r tcoefs3}
rownames(coefs)
print(res, col.names = TRUE)
```

* Beta coefficients path

```{r selected-betas}
len.max <- 7
whch_1se <- which(sel == idx_1se)
whch_min <- which(sel == idx_min)
selx <- sel[whch_1se:whch_min]
selxx <- selx
if (length(selx) > len.max) selxx <- c(selx[1:(len.max-1)], selx[length(selx)])
c(idx_1se, idx_min)

resx <- coefs[, selxx]
round(resx,4)
```


