---
title: 'Cox Models multiple -preamble'
author: "Chunyi Wu and Andrzej Galecki"
date: "`r format(Sys.Date())`"
output:
  rmdformats::readthedown:
    lightbox: true
    use_bookdown: true
params:
   alphax: -99
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```

```{r utilsag-version-2test, include = FALSE}
uag_ver0 <- "0.2.1" # utilsag version tested to execute this script
```

```{r install-tested-utilsag-child, child = "_install-tested-utilsag.Rmd", include = FALSE}
```

This script was executed using `utilsag` package stored in Github branch *`r uag_ref0`*,
as requested.


```{r utilsag-lodaed}
library(utilsag)
```

```{r params-read}
alphax <- params$alphax
```



<!-- ========== Read raw data   ================= -->

# Read raw data (alpha = `r alphax`)

```{r  Read-raw-data}
excel_datain <- "./Data/data_example.xlsx" # Enter dataset name
dt <- read_excel(excel_datain)
```

## Data prep. 

* Data are prepared for Cox regression.        
```{r data-prep}
 dt <- dt  %>% rename(time = FU_TIME, status = CASE_CONTROL) %>% 
  mutate(log10_DU_ACR = log10(DU_ACR))  %>% filter(time>0) %>% filter(BL_eGFR >= 45)

dim(dt) # Number of rows and columns in the input data
```


## Variables

```{r Variables-used-Cox-model}
prot_npx <- c("KIM1.npx","SYND1.npx","IL.1RT1.npx",   "WFDC2.npx", "CD27.npx",
              "TNFRSF10A.npx","LAYN.npx","PVRL4.npx", "EDA2R.npx","TNFRSF4.npx",
              "GFR_alpha_1_npx","TNF.R1.npx","PI3_npx", "EFNA4.npx","TNF.R2.npx" ,
              "DLL1.npx", "TNFRSF6B.npx", "CD160.npx", "EPHA2.npx","RELT.npx",
              "LTBR.npx") 
surv_vars <-  c("time", "status")
clin_vars <- c("BL_eGFR","B_HBA1C_PRC","log10_DU_ACR")  ## excluded,"SEX","AGE_TL")
xvars <- c(clin_vars, prot_npx)
vars <- c(surv_vars, prot_npx, clin_vars) # For Model M2
clin_vars
```


## Create dataset `datax`

* Create dataset `datax` that contains subset of all vars needed for the analysis.

```{r data-prep1}
datax <- dt  %>% select(all_of(vars)) 
dim(datax)         
```


## Complete cases

* Keep rows with complete cases in `train_data` dataframe.


```{r data-prep2}
#drop any records with NA
train_data <- datax %>% drop_na()    
```

## survSplit-cut

```{r survSplit-cut}
survSplit_cut <- 10  # Time horizon 10 years
```

## Create `train_data15`

* Create data that is censored at `r survSplit_cut` years (prediction horizon)
* extra varibles are created

```{r survSplit}
temp <- survSplit(Surv(time,  status) ~ ., data = train_data, cut = survSplit_cut,
                  episode = "epsd")
dim(train_data)

train_data15 <- subset(temp, epsd == 1)  # only the first ?? years
dim(train_data15)
names(train_data15)
train_data_saved    <- train_data15
```

## Selected variables

```{r head15}
range(train_data15$time)
range(train_data15$epsd)
tt <- train_data15 %>% select(BL_eGFR, tstart, time, status, epsd)
# head(tt)
# tail(tt)
```

## X matrix

* Use `train_data15` to create X matrix and Surv object required by `glmnet`.

```{r model-mtx}
# Create X matrix
# -dtx <- subset(train_data, select=-c(time,status))
dtx <- train_data15 %>% select(all_of(xvars)) 
x <- model.matrix(~0 +., data=dtx)
dim(x)
colnames(x)
```

## Create Surv object

```{r surv-object}
y <- data.matrix(train_data15[,c("time", "status")])
ySurv <- survival::Surv(y[,"time"], y[, "status"])
```

## Create cv-folds 

* (nfolds = 10 is default in `cv.glmnet`)

```{r foldid-1}
set.seed(1234)
nfolds <-10
foldid <- sample(1:nfolds, size = nrow(train_data15), replace=TRUE)
length(foldid)
```

* Table with fold sizes is included below.

```{r foldid-2}
tibble(foldid = foldid) %>% group_by(foldid) %>% count()
```

#  Elastic net


* Elastic net with penalty.factor
* Clinical covariates forced into the model (pen =0)
* Selected proteins with increased penalty factor

## glmnet

* All predictors included 

```{r beta1}
clin_vars
col_nmsx <- colnames(x)
ncolx <- length(colnames(x))
pen <- rep(1, times= ncolx) # default penalty =1
names(pen) <- col_nmsx
pen[clin_vars] <- 0
# penalized proteins
prot_pen <- c("TNFRSF10A.npx", "LAYN.npx", "EDA2R.npx", "TNFRSF4.npx", "TNFRSF6B.npx", "EPHA2.npx")
pen[prot_pen] <- 5
pen
fit1 <- glmnet(x, ySurv, family = "cox", alpha=alphax, penalty.factor = pen)
(gl_fit1 <- myglance(fit1))
(td_fit1 <- mytidy(fit1))

plot(fit1)
## plot(fit1, xvar = "lambda", label = TRUE)

```

```{r  ggplot1-coefs}
dfg1 <- td_fit1 %>% unnest(beta) %>% group_by(step) 

ggplot(dfg1 , aes(x = log(lambda), y = estimate)) +
       geom_line(aes(color = term)) + ylab("Coefficients")
```


## cv.glmnet

```{r cv-fit1}
cvfit1 <- cv.glmnet(x, ySurv, family = "cox", alpha=alphax, penalty.factor = pen, foldid =foldid)
cvfit1 # saved


(mygl1 <- myglance(cvfit1))
lmbda <- mygl1[,c("lambda.min", "lambda.1se")] %>% 
           mutate(log_lmin = log(lambda.min), log_l1se = log(lambda.1se))
lmbda

#mytidy(cvfit1)
plot(cvfit1)
```

* C_V using Cindex

```{r cvfit-plot-Cindex}
cvfit1_Cindex <- cv.glmnet(x, ySurv, family = "cox",  type.measure ="C",
                  alpha=alphax, penalty.factor = pen, foldid =foldid)
plot(cvfit1_Cindex) # saved

```

## c-v partial likelihood dev

* cross-validated  partial likelihood deviance


```{r td-cvfit1}
td_cvfit1_init <- mytidy_Surv(cvfit1, x, ySurv)
td_cvfit1_init$comment <- paste0("alpha=", alphax)
td_cvfit1 <- td_cvfit1_init %>% mutate(alpha =alphax)%>%
   relocate(alpha, .before = step)
td_cv1 <-td_cvfit1  %>% select(-c(beta, conf.low, conf.high, df, AICc)) 
td_cv1 %>% filter(mincv != "-") %>% print(n = 500)

```

## beta coefficients

```{r beta-coefs}
beta1_init <- td_cvfit1 %>% select(c(step,lambda, nzero, beta, mincv, comment)) %>% 
                unnest(beta) %>% group_by(step) 

beta1 <- beta1_init %>% mutate(alpha = alphax) %>% relocate(alpha, .before = step)
nrow(beta1)               
beta1 %>% filter(mincv != "-") %>% print(n=400)
```

```{r coef-beta}
bx1 <- beta1 %>% select(term, estimate, mincv)
bx1 %>% filter(mincv == "1se" | mincv == "min" | mincv == "+")
```

## Coefficients path

* Plot using glmnet fit object


```{r beta-path}
plot(fit1, xvar = "lambda", label = TRUE)
```



* Template for beta coefficients.
* Rows correspond to (descending) sequence of lambda hyperparameters
* Columns correspond to covariates

```{r tcoefs1}
coefs <- coef(fit1)
ncoefs <- nrow(coefs)
coef_nms <- rownames(coefs)
snms <- colnames(coefs)
idx_1se <- mygl1 %>% select(index_1se) %>% pull()
idx_min <- mygl1 %>% select(index_min) %>% pull()

snms[idx_1se] <- paste0("<" ,snms[idx_1se])
snms[idx_min] <- paste0(snms[idx_min], ">")


tx <- t(coefs)
ttx <- t(tx)
tx0 <- t(ttx != 0)
idx <- apply(tx0, 1, function(x) paste(ifelse(x, 1, 0), collapse =""))
idxr <- rank(idx)
tmp <- idxr[c(idx_1se, idx_min)]
idxr[c(idx_1se, idx_min)] <- tmp + 0.01

sel <- which(!duplicated(idxr)) # idx -> idxr
res <- tx0[sel,]
rownames(res) <- snms[sel]
```

```{r tcoefs2}
cnms <- 1:ncoefs
colnames(res) <- cnms
```
* Predictors 

```{r tcoefs3}
tibble(x=rownames(coefs), pen=pen) %>% print (n=100)

print(res, col.names = TRUE)
```

* Beta coefficients path

```{r selected-betas}
len.max <- 7
whch_1se <- which(sel == idx_1se)
whch_min <- which(sel == idx_min)
selx <- sel[whch_1se:whch_min]
selxx <- selx
if (length(selx) > len.max) selxx <- c(selx[1:(len.max-1)], selx[length(selx)])
c(idx_1se, idx_min)
resx <- coefs[, selxx]
round(resx,4)
```

```{r rmo}
rm(idx_1se, idx_min, tmp, whch_1se, whch_min)
#rm(coefs,  resx)

```


