---
title:  "`r paste0(bnm,'.Rmd:')`"
author: "`r username`"
date: "`r format(Sys.Date())`"
output:
  rmdformats::readthedown:
    lightbox: true
    use_bookdown: true
---

<!-- SETUP  -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```

```{r setup-npx-data, include=FALSE}

source("./Rfun/summarize_variable.R")
source("./Rfun/summarize_data.R")
source("./Rfun/check_survival_consistency2.R")
source("20.setup_npxdata.txt")

```

<!-- Section: Original NPX data -->

Script `r bnm` executed by `r username`.

# NPX Data: `npx_data_orig`

* Original NPX Data path: `r npx_dtpath`

* `r model_desc`


```{r orig-summary-stats, echo=FALSE}
summary_stats = summarize_data(npx_data_orig)
knitr::kable(summary_stats, caption = "Summary Statistics for original NPX data")
```


Script `r bnm` executed by `r username`.

* Original NPX Data path: `r npx_dtpath`

* `r model_desc`

* Filter BL_eGFR >= 45
* Vars renamed: time = FU_TIME, status = CASE_CONTROL, time_death = FU_TIME_Death, death = Death

<!-- Check surv vars -->

## Check time vars(1)

* Checking for missing values in time vars. Auxiliary data `tvars_sel` should be empty 
```{r check-surv-vars-orig}
tvars_sel = npx_data_orig %>% 
  select(c(INDEX,FU_TIME,  CASE_CONTROL, FU_TIME_Death, Death))
sel = complete.cases(tvars_sel)
head(tvars_sel[!sel, ])
```
## Check time vars(2)


* Check time vars in `npx_data_orig` data for consistency

```{r check2-npx_data_orig}

miss_res1 =check_survival_consistency2( npx_data_orig, 
                          index_var = "INDEX", 
                          time_var = "FU_TIME", 
                          status_var = "CASE_CONTROL", 
                          time_death_var = "FU_TIME_Death", 
                          death_var = "Death")
#summ1_list= miss_res1$summary
#miss_summ1 <- data.frame(Value = unlist(summ1_list))
#miss_summ1
#miss_res1$issues
miss_res1$problematic_rows
```

<!--  Section: NPX `npx_dt2` data -->

# `npx_dt2` auxiliary data

* Filter BL_eGFR >= 45
* Vars renamed: time = FU_TIME, status = CASE_CONTROL, time_death = FU_TIME_Death, death = Death
* time vars only 

```{r npx-dt-summary-stats-npx-dt, echo=FALSE}
npx_dt2 = npx_dt %>% select(-c(time_death, death,event_time,event_type, composite_time, composite_status ))
summary_stats = summarize_data(npx_dt2)
knitr::kable(summary_stats, caption = "Summary Statistics for npx_dt2 data")
```

<!--  Section: NPX npxdata_all data -->

# `npxdata_all` data

* Filter BL_eGFR >= 45
* Vars renamed: time = FU_TIME, status = CASE_CONTROL, time_death = FU_TIME_Death, death = Death
* Var: time_death if missing then replaced with `r paste0("survSplit_cut=",survSplit_cut)` 
* Vars:  created event_time,event_type, composite_time, composite_type

## Summary

```{r npx-dt-summary-stats-npx-dt_all, echo=FALSE}
table(status = npxdata_all$status)
table(death = npxdata_all$death)
table(composite_status = npxdata_all$composite_status)
table(event_tye = npxdata_all$event_type)

summary_stats = summarize_data(npxdata_all)
knitr::kable(summary_stats, caption = "Summary Statistics for npxdata_all data")
```

## Time vars check

* Check time vars in `npxdata_all` data for consistency

```{r check2-npxdata_all}
                          
miss_res2 =check_survival_consistency2(data=npxdata_all, 
                          index_var = "INDEX", 
                          time_var = "time", 
                          status_var = "status", 
                          time_death_var = "time_death", 
                          death_var = "death")
#summ2_list= miss_res2$summary
#miss_summ2 <- data.frame(Value = unlist(summ2_list))
#miss_summ2

#miss_res2$issues
miss_res2$problematic_rows
```


<!-- Time vars -->

```{r tvars_dt}
if (username == "agalecki"){
  tvars_dt0 = npxdata_all %>% select(c(INDEX,time, status, time_death, death, event_time,event_type, composite_time, 
          composite_status))
  tvars_dt = tvars_dt0 %>% rename(tmdth = time_death, etime= event_time, dth =death, cmp_tm = composite_time, cmp_stt= composite_status,
     stt=status, etp = event_type)       
  cols_to_round <- c("time", "tmdth", "etime","cmp_tm")
  tvars_dt[, cols_to_round] <- lapply(tvars_dt[, cols_to_round], round, digits = 2)

  print(tvars_dt)
}
```