---
title:  "`r paste0(bnm,'.Rmd:')`"
author: "`r username`"
date: "`r format(Sys.Date())`"
output:
  rmdformats::readthedown:
    lightbox: true
    use_bookdown: true
---

<!-- SETUP  -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```

```{r setup-npx-data, include=FALSE}

#source("./Rfun/summarize_variable.R")
#source("./Rfun/summarize_data.R")
#source("./Rfun/check_survival_consistency2.R")
#source("./Rfun/get_newdata_linear_predictor.R")
## source("./Rfun/calculate_survival_probabilities.R")
# source("./Rfun/crr_fg_train.R")
source("./Rfun/calculate_brier_score_cvfit.R")
source("./Rfun/calibration_analysis.R")
source("./Rfun/calibration_table_with_diagnostics.R")
source("./Rfun/compute_nri_idi_single.R")
source("./Rfun/cox_glmnet_train.R")
source("20.setup_npxdata.txt")
load("./validation_input/17-cox-0.25M3.Rdata", verbose=TRUE)
cvfit = res_cvfit[["ALLx"]] # cv.glmnet class

```

Return objects with baseline_cumulative hazards

* Cox with clinical predictors only

```{r create-cum-hazard1-coxph-clin_vars3-test}

# Define predictors and survival variables
surv_vars <- c("time", "status")
all_vars <- c(surv_vars, clin_vars3)

# Subset npxdata_all
data_surv <- npxdata_all %>% select(all_of(all_vars))
coxph_fit_clin3 <- coxph(Surv(time, status) ~ ., data = data_surv)  # [, clin_vars3])
bas_cumhazdf <- basehaz(coxph_fit_clin3, centered = FALSE)
min_haz <- min(bas_cumhazdf$hazard[bas_cumhazdf$hazard > 0]) # Fix zeros, if any
bas_cumhazdf$hazard[bas_cumhazdf$hazard == 0] <- min_haz
bas_cumhazdf$hazard <- pmin(bas_cumhazdf$hazard, quantile(bas_cumhazdf$hazard, 0.99)) # Cap outliers
bas_cumhazdf_clin3  <- bas_cumhazdf
# rm(bas_cumhazdf_clin3)

coxph_clin3_mdl <-
  list(fit = coxph_fit_clin3,
       cumBasHazdf =  bas_cumhazdf)
```



* `cvfit` loaded
* create baseline hazard

```{r create-cum-hazard1-cv.glmnet-M3-test}

cvfit_predictors =c(prot_npx, clin_vars3) # 21 + 3 
surv_vars <- c("time", "status")
cvfit_predictors <- c(prot_npx, clin_vars3)  # 21 proteomic + 3 clinical
table(npxdata_all$status)  # Check distribution of 0, 1
pen  = cvfit_auxArgs$penalty.factor
foldid = cvfit_auxArgs$foldid

cox_glmnet_result <- cox_glmnet_train(predictors = cvfit_predictors, 
                                      penalty_factor = pen,
                                      foldid = foldid,
                                      verbose = TRUE)
str(cox_glmnet_result)
```

* `cox_glmnet_result` list contains the following components:
cvfit_scaled, bas_cumhazdf, lin_pred, scaling_params 


```{r crr-clin-vars3,eval=FALSE, echo=FALSE}
# CHUNK SKIPPED
table(npxdata_all$event_type)  # Check distribution of 0, 1, 2
bas_cumhazdf_crr_clin3 <- crr_fg_train(clin_vars3)
str(bas_cumhazdf_crr_clin3)

```

# Brier score

```{r brier-score}
eval_time <- 10  # Evaluate Brier score at t=10
brier_score <- calculate_brier_score(cox_glmnet_result = cox_glmnet_result, 
                                     data = npxdata_all, 
                                     eval_time = eval_time, 
                                     event_vars = c("time", "status"))
cat("Brier score at time", eval_time, ":", brier_score, "\n")
```

##  Diagnostic output
* Diagnostic output for brier sore

```{r brier-Diagnostic-output}
lin_pred <- cox_glmnet_result$lin_pred
surv_prob <- exp(-approx(cox_glmnet_result$bas_cumhazdf$time, 
                         cox_glmnet_result$bas_cumhazdf$hazard, 
                         xout = eval_time, method = "linear")$y * exp(lin_pred))
cat("Range of lin_pred:", range(lin_pred, na.rm = TRUE), "\n")
cat("Number of unique survival probabilities:", length(unique(surv_prob)), "\n")
summary(surv_prob)

```

# Calibration

## Calibration (excld = TRUE)

### Analysis(1)

```{r calibration-analysis}
# Example usage
eval_time <- 10
tt1 <- calibration_analysis(cox_glmnet_result = cox_glmnet_result, 
                           data = npxdata_all, 
                           eval_time = eval_time, 
                           event_vars = c("time", "status"), 
                           n_bins = 10, 
                           clip_value = 1e-3, 
                           exclude_extremes = TRUE)
cat("Calibration Slope:", tt1$calib_slope, "\n")
cat("Calibration Intercept:", tt1$calib_intercept, "\n")
cat("Standard Error (Slope):", tt1$se_slope, "\n")
cat("Standard Error (Intercept):", tt1$se_intercept, "\n")
cat("Slope 95% CI:", tt1$ci_slope[1], "to", tt1$ci_slope[2], "\n")
cat("Intercept 95% CI:", tt1$ci_intercept[1], "to", tt1$ci_intercept[2], "\n")
if (!is.null(tt1$calibration_data)) {
  print(tt1$calibration_data)
}
# plot(calib_data$mean_pred, calib_data$obs_surv ...

```
### Diagnostics Table (1)

Example usage with calibration_analysis output:

```{r calibration-table-with-diagnostics} 
eval_time <- 10
table_result1 <- calibration_table_with_diagnostics(
  calib_result = tt1, 
  cox_glmnet_result = cox_glmnet_result, 
  data = npxdata_all, 
  eval_time = eval_time, 
  event_vars = c("time", "status")
)

# Print the table again for reference
cat("\nCalibration Table:\n")
print(table_result1$calibration_table)
#rm(tt1,table_result1)

```

## Calibration (excld = FALSE)

### Analysis(2)

```{r calibration-analysis2}
# Example usage
eval_time <- 10
tt2 <- calibration_analysis(cox_glmnet_result = cox_glmnet_result, 
                           data = npxdata_all, 
                           eval_time = eval_time, 
                           event_vars = c("time", "status"), 
                           n_bins = 10, 
                           clip_value = 1e-3, 
                           exclude_extremes = FALSE)
cat("Calibration Slope:", tt2$calib_slope, "\n")
cat("Calibration Intercept:", tt2$calib_intercept, "\n")
cat("Standard Error (Slope):", tt2$se_slope, "\n")
cat("Standard Error (Intercept):", tt2$se_intercept, "\n")
cat("Slope 95% CI:", tt2$ci_slope[1], "to", tt2$ci_slope[2], "\n")
cat("Intercept 95% CI:", tt2$ci_intercept[1], "to", tt2$ci_intercept[2], "\n")
if (!is.null(tt2$calibration_data)) {
  print(tt2$calibration_data)
}
# plot(calib_data$mean_pred, calib_data$obs_surv ...

```
### Diagnostics Table (2)

Example usage with calibration_analysis output:

```{r calibration-table-with-diagnostics2} 
eval_time <- 10
table_result2 <- calibration_table_with_diagnostics(
  calib_result = tt2, 
  cox_glmnet_result = cox_glmnet_result, 
  data = npxdata_all, 
  eval_time = eval_time, 
  event_vars = c("time", "status")
)

# Print the table again for reference
cat("\nCalibration Table:\n")
print(table_result2$calibration_table)
# rm(tt2,table_result2)

```

# NRI/IDI


## Single (excld=TRUE)

```{r IDI-NRI1}
recalb1 = list(slope = tt1$calib_slope, intercept= tt1$calib_intercept)
recalb1
single_result1 <- compute_nri_idi_single(
  new_model_result = cox_glmnet_result,
  clin_vars = clin_vars3,
  biomarker_vars = prot_npx, # biomarkers,
  data = npxdata_all,
  eval_time = eval_time,
  event_vars = c("time", "status"),
  recalibrate = recalb1    # list(slope = 2.444079, intercept = 1.756214)
 )

```

## Single (excld=FALSE)

```{r IDI-NRI2}
recalb2 = list(slope = tt2$calib_slope, intercept= tt2$calib_intercept)
recalb2
single_result2 <- compute_nri_idi_single(
  new_model_result = cox_glmnet_result,
  clin_vars = clin_vars3,
  biomarker_vars = prot_npx, # biomarkers,
  data = npxdata_all,
  eval_time = eval_time,
  event_vars = c("time", "status"),
  recalibrate = recalb2    # list(slope = 2.444079, intercept = 1.756214)
 )
#cat("IDI:", idi_result$point$IDI[1,1], "\n")
#cat("NRI:", idi_result$point$TXI[1], "\n")
# Removed: print(idi_result)
# idi_res2 = single_result2$idi_result
# str(idi_res2)

```

# Save

```{r save-results}
onms = c("tt1", "table_result1",  "tt2", "table_result2", "single_result1",
    "single_result2")
outpath = paste0("./validation_save/22.cvfit2test.Rdata")
save(file=outpath, list = onms)
```
