---
title: 'Validation of Cox Model M2'
author: "Chunyi Wu and Andrzej Galecki"
date: "`r as.character(Sys.Date(), format = '%A %B %d, %Y')`"
output:
  rmdformats::readthedown:
    lightbox: true
    use_bookdown: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```

```{r data, include = FALSE}
if (!require('pacman')) install.packages('pacman', repos = "http://cran.us.r-project.org")
library(pacman)

pacman::p_load(
  readxl,       # load excel file
  glmnet,       # Lasso and Elastic-Net Regularized Generalized Linear Models
  rmdformats,   # rmd formats
  rmarkdown,    # rmarkdown
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics 
  janitor,      # adding totals and percents to tables
  flextable,    # format the output table
  gtsummary,    # calculate summary statistics and format the results
  sjPlot,       # correlation matrix
  purrr,        # enhances R functional programming (FP) toolki 
  tidyr,        #Tools to help to create tidy data
  ggplot2,      #Plot the results
  glmnetUtils,  #Glmnet models for multiple alpha
  coefplot,     # Plotting Model Coefficients
  survival,     #survival model 
  tidymodels,   #for modeling and machine learning using tidyverse principles
  survivalROC   #survivalROC
  )
```

```{r utilsag-version-2test, include = FALSE}
uag_ver0 <- "0.2.1" # utilsag version tested to execute this script
```

```{r install-tested-utilsag-child, child = "_install-tested-utilsag.Rmd", include = FALSE}
```

This script was executed using `utilsag` package stored in Github branch *`r uag_ref0`*,
as requested.


```{r utilsag-lodaed}
library(utilsag)
mod_lbl <-"M3"
alphax <- 0.25
survSplit_cut <- 10

```



# Introduction (`r mod_lbl`)

In this report we consider Cox regression model `r mod_lbl` for the time-to-event `time` variable. 

<!-- Conditonal text inserted -->  
`r if(mod_lbl == "M0"){
"Model M0 contains baseline HbA1c, log10(ACR), BL_eGFR, SEX, and AGE_TL(Baseline Age) as candidate predictors"
}`

`r if(mod_lbl == "M1"){
"Model M1 contains 21 proteins as candidate predictors."
}`


`r if(mod_lbl == "M2"){
"Model M2 contains 21 proteins and Baseline HbA1c, log10(ACR), BL_eGFR, SEX, and AGE_TL(Baseline Age) as candidate predictors"
}`
<!-- Conditonal text ends --> 

Notes: 

* Originally `time` variable was named `FU_TIME`.  
* `status` (0/1) variable, originally named `CASE_CONTROL`, is coded 0 for Controls (i.e., ESKD event has not occurred), and 1 for Cases (i.e., event has occurred)


## Test data prep

Test data prepared for validation of Cox regression model `r mod_lbl`.        

```{r, prepare-data, echo=TRUE, message=FALSE, warning=FALSE}
#read the data that is stored under the `validation_input` folder

dt <- read_excel("./validation_input/data_validation.xlsx")

dt <- dt  %>% rename(time = FU_TIME, status = CASE_CONTROL) %>% 
  mutate(log10_DU_ACR = log10(DU_ACR))  %>% filter(time>0) %>% filter(BL_eGFR >= 45)

dim(dt) # Number of rows and columns in the input data
```
Variable names used in Cox regression model are stored in the following vectors:

* Proteins in `prot_npx`,
* Time/status in `surv_vars`,
* Clinical variables in `clin_vars`

```{r Variables-used-Cox}
prot_npx <- c("KIM1.npx","SYND1.npx","IL.1RT1.npx",   "WFDC2.npx", "CD27.npx",
              "TNFRSF10A.npx","LAYN.npx","PVRL4.npx", "EDA2R.npx","TNFRSF4.npx",
              "GFR_alpha_1_npx","TNF.R1.npx","PI3_npx", "EFNA4.npx","TNF.R2.npx" ,
              "DLL1.npx", "TNFRSF6B.npx", "CD160.npx", "EPHA2.npx","RELT.npx",
              "LTBR.npx") 
surv_vars <-  c("time", "status")
clin_vars <- c("BL_eGFR","B_HBA1C_PRC","log10_DU_ACR","SEX","AGE_TL")
clin_vars3 <-  c("BL_eGFR","B_HBA1C_PRC","log10_DU_ACR")
xvars <- c(clin_vars3, prot_npx)
```



Data preparation: 

* Vector `vars`  contains all variable names needed to specify model `r mod_lbl` 
considered in this document

```{r select-mvars, include =FALSE}

if (mod_lbl == "M0") mvars <- c(surv_vars, clin_vars) 
if (mod_lbl == "M1") mvars <- c(surv_vars, prot_npx, clin_vars) 
if (mod_lbl == "M2") mvars <- c(surv_vars, prot_npx, clin_vars) 
if (mod_lbl == "M3") mvars <- c(surv_vars, prot_npx, clin_vars3) 

```

```{r vars vector}
mvars
```

* Create dataset `data` that contains subset of all vars needed for the analysis.

```{r data-prep1}
data <- dt  %>% select(all_of(mvars)) 
dim(data)         
```

* Keep rows with complete cases in `test_data` dataframe.


```{r data-prep2}
# drop any records with NA
test_data <- data %>% drop_na()    

# Total sample size
nrow(test_data)
```

# Preparatory steps

Load fles with info on selected models fitted using traininig data.

```{r load-train-objects}
fpath1 <- paste0("./validation_input/17-cox-0.25", mod_lbl, ".Rdata")
load (file = fpath1, verbose = TRUE)
#- fpath2 <- paste0("./save/5.Cox_tidy2_", mod_lbl, ".Rdata")
# save(mod_selected, cva_pfit, mod_all, sel3long, file = fpath2)

```

In this document, we assess performance of the model `r mod_lbl` with the hyperparameters
listed below:

```{r hyperparametrs-selected}
alphax
```


## survSplit

* Create test_data  that is administratively censored at `r survSplit_cut`(time horizon)

```{r survSplit}
message("survSplit_cut =", survSplit_cut)

temp <- survSplit(Surv(time,  status) ~ ., data = test_data, cut = survSplit_cut,
                  episode = "epsd")
dim(test_data)
test_data15 <- subset(temp, epsd == 1)  # only the first ?? years
dim(test_data15)
test_data_saved <- test_data
test_data    <-  test_data15
```



Matrix X and `Surv` object extracted from the `test_data`

```{r x-surv-test}
# - dtx_test <- subset(test_data, select=-c(time,status))
dtx_test <- test_data %>% select(all_of(xvars)) 
x_test <- model.matrix(~0 +., data=dtx_test)
dim(x_test)
colnames(x_test)
# Create Surv object
y_test <- data.matrix(test_data[,c("time", "status")])
ySurv_test <- survival::Surv(y_test[,"time"], y_test[, "status"])
```

We extract model fit for a given $\alpha$ and store it in `cv_opt_model` object
for later use.

```{r cva-glmnet-fit}
names(res_cvfit)
cvfit <- res_cvfit$ALLx

pfit_aopt <- cvfit$glmnet
class(pfit_aopt)
names(pfit_aopt)
pfit_aopt$lambda

lmbda <- pfit_aopt$lambda
lmbda_sel <-lmbda[c(1, 15, 18,24,30)]
lmbda_opt <- lmbda_sel[1]
```

# Model Performance 




Predictive performance of a selected model using _standard_ approach.


## C-index

Step 1: Linear predictor estimation 

Estimate linear predictor values for the Cox model fitted to test data. 

```{r calc-pred2}
predM_lpmtx <- predict(pfit_aopt, newx = x_test, type = "link") # Matrix
predM_lp <- as.vector(predict(pfit_aopt, newx = x_test, type = "link", s = lmbda_opt))
# predM_lp <- - predM_lp # !!!
predM_lp[1:10]
summary(predM_lp)
```
 

Step2: C-index calculations

Ref: Harrel Jr, F. E. and Lee, K. L. and Mark, D. B. (1996) Tutorial in biostatistics:
multivariable prognostic models: issues in developing models, evaluating assumptions 
and adequacy, and measuring and reducing error, _Statistics in Medicine_, 15, pages 361-387.


```{r C-index-Mod}
Cindex_lmbda <- apply(predM_lpmtx, 2, Cindex, y = ySurv_test) # Multiple lambda
length(Cindex_lmbda)
Cindex_lmbda[1:12] 

Cindex(predM_lp, ySurv_test)   # For optimal lambda
```

## Time-dependent ROC

* source: (https://datascienceplus.com/time-dependent-roc-for-survival-prediction-models-in-r)


```{r surv-roc}
dataM_test <- test_data
                
## Augment `dataM_test` with linear predictor
 dataM_test$predM_lp <- predM_lp

# Evaluate every 2.5 years
 
tx <- 2.5* c(1,2,3,4) # ROC time points
tx
predM_lp[1:10]
survROC_lp <- create_survivalROC_data(tx, dataM_test, predM_lp)
 
survROC_lp %>% print (n=50)
```

* Plot Time-dependent ROC every 2.5 years

```{r surv-roc-plot}
## Plot Time-dependent ROC every 2.5 years
plot_timedep_ROC(survROC_lp) 
```


```{r exit0, include=FALSE}
knitr::knit_exit()
```
