---
title: 'Validation of Cox Model M2'
author: "Chunyi Wu and Andrzej Galecki"
date: "`r format(Sys.Date())`"
output:
  rmdformats::readthedown:
    lightbox: true
    use_bookdown: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```

```{r data, include = FALSE}
if (!require('pacman')) install.packages('pacman', repos = "http://cran.us.r-project.org")
library(pacman)

pacman::p_load(
  readxl,       # load excel file
  glmnet,       # Lasso and Elastic-Net Regularized Generalized Linear Models
  rmdformats,   # rmd formats
  rmarkdown,    # rmarkdown
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics 
  janitor,      # adding totals and percents to tables
  flextable,    # format the output table
  gtsummary,    # calculate summary statistics and format the results
  sjPlot,       # correlation matrix
  purrr,        # enhances R functional programming (FP) toolki 
  tidyr,        #Tools to help to create tidy data
  ggplot2,      #Plot the results
  glmnetUtils,  #Glmnet models for multiple alpha
  coefplot,     # Plotting Model Coefficients
  survival,     #survival model 
  tidymodels,   #for modeling and machine learning using tidyverse principles
  survivalROC   #survivalROC
  )
```

```{r utilsag-version-2test, include = FALSE}
uag_ver0 <- "0.2.1" # utilsag version tested to execute this script
```

```{r install-tested-utilsag-child, child = "_install-tested-utilsag.Rmd", include = FALSE}
```

This script was executed using `utilsag` package stored in Github branch *`r uag_ref0`*,
as requested.


```{r utilsag-lodaed}
library(utilsag)
mod_lbl <-"M3"
alphax <- 0.25
survSplit_cut <- 10

```



# Introduction (`r mod_lbl`)

In this report we consider Cox regression model `r mod_lbl` for the time-to-event `time` variable. 

<!-- Conditonal text inserted -->  
`r if(mod_lbl == "M0"){
"Model M0 contains baseline HbA1c, log10(ACR), BL_eGFR, SEX, and AGE_TL(Baseline Age) as candidate predictors"
}`

`r if(mod_lbl == "M1"){
"Model M1 contains 21 proteins as candidate predictors."
}`


`r if(mod_lbl == "M2"){
"Model M2 contains 21 proteins and Baseline HbA1c, log10(ACR), BL_eGFR, SEX, and AGE_TL(Baseline Age) as candidate predictors"
}`

`r if(mod_lbl == "M3"){
"Model M3 contains 21 proteins and Baseline HbA1c, log10(ACR), BL_eGFR as candidate predictors. Clinical covariates
are forced into the model."
}`

<!-- Conditonal text ends --> 

Notes: 

* Originally `time` variable was named `FU_TIME`.  
* `status` (0/1) variable, originally named `CASE_CONTROL`, is coded 0 for Controls (i.e., ESKD event has not occurred), and 1 for Cases (i.e., event has occurred)


## Test data prep

Test data prepared for validation of Cox regression model `r mod_lbl`.        

```{r, prepare-data, echo=TRUE, message=FALSE, warning=FALSE}
#read the data that is stored under the `validation_input` folder

# Change the filename below, if needed
dt <- read_excel("./validation_input/data_validation.xlsx")

dt <- dt  %>% rename(time = FU_TIME, status = CASE_CONTROL) %>% 
  mutate(log10_DU_ACR = log10(DU_ACR))  %>% filter(time>0) %>% filter(BL_eGFR >= 45)

dim(dt) # Number of rows and columns in the input data
```
Variable names used in Cox regression model are stored in the following vectors:

* Proteins in `prot_npx`,
* Time/status in `surv_vars`,
* Clinical variables in `clin_vars`

```{r Variables-used-Cox}
prot_npx <- c("KIM1.npx","SYND1.npx","IL.1RT1.npx",   "WFDC2.npx", "CD27.npx",
              "TNFRSF10A.npx","LAYN.npx","PVRL4.npx", "EDA2R.npx","TNFRSF4.npx",
              "GFR_alpha_1_npx","TNF.R1.npx","PI3_npx", "EFNA4.npx","TNF.R2.npx" ,
              "DLL1.npx", "TNFRSF6B.npx", "CD160.npx", "EPHA2.npx","RELT.npx",
              "LTBR.npx") 
surv_vars <-  c("time", "status")
clin_vars <- c("BL_eGFR","B_HBA1C_PRC","log10_DU_ACR","SEX","AGE_TL")
clin_vars3 <-  c("BL_eGFR","B_HBA1C_PRC","log10_DU_ACR")
xvars <- c(clin_vars3, prot_npx)
```



Data preparation: 

* Vector `vars`  contains all variable names needed to specify model `r mod_lbl` 
considered in this document

```{r select-mvars, include =FALSE}

if (mod_lbl == "M0") mvars <- c(surv_vars, clin_vars) 
if (mod_lbl == "M1") mvars <- c(surv_vars, prot_npx, clin_vars) 
if (mod_lbl == "M2") mvars <- c(surv_vars, prot_npx, clin_vars) 
if (mod_lbl == "M3") mvars <- c(surv_vars, prot_npx, clin_vars3) 

```

```{r vars vector}
mvars
```

* Create dataset `data` that contains subset of all vars needed for the analysis.

```{r data-prep1}
data <- dt  %>% select(all_of(mvars)) 
dim(data)         
```

* Keep rows with complete cases in `test_data` dataframe.


```{r data-prep2}
# drop any records with NA
test_data <- data %>% drop_na()    

# Total sample size
nrow(test_data)
```

# Preparatory steps

Load fles with info on selected models fitted using traininig data.

```{r load-train-objects}
fpath1 <- paste0("./validation_input/17-cox-0.25", mod_lbl, ".Rdata")
load (file = fpath1, verbose = TRUE)
#- fpath2 <- paste0("./save/5.Cox_tidy2_", mod_lbl, ".Rdata")
# save(mod_selected, cva_pfit, mod_all, sel3long, file = fpath2)

```

In this document, we assess performance of the model `r mod_lbl` with the hyperparameters
listed below:

```{r hyperparametrs-selected}
alphax
```


## survSplit

* Create test_data  that is administratively censored at `r survSplit_cut`(time horizon)

```{r survSplit}
message("survSplit_cut =", survSplit_cut)

temp <- survSplit(Surv(time,  status) ~ ., data = test_data, cut = survSplit_cut,
                  episode = "epsd")
dim(test_data)
test_data15 <- subset(temp, epsd == 1)  # only the first ?? years
dim(test_data15)
test_data_saved <- test_data
test_data    <-  test_data15
```



Matrix X and `Surv` object extracted from the `test_data`

```{r x-surv-test}
# - dtx_test <- subset(test_data, select=-c(time,status))
dtx_test <- test_data %>% select(all_of(xvars)) 
x_test <- model.matrix(~0 +., data=dtx_test)
dim(x_test)
colnames(x_test)
# Create Surv object
y_test <- data.matrix(test_data[,c("time", "status")])
ySurv_test <- survival::Surv(y_test[,"time"], y_test[, "status"])
```

Auxiliary calculations:

* cens10: Censored obs. Follow-up less than 10 years and event has not occurred
* Cp10_test: Percent of censored obs (based on cens10)
* PN10_test: PN is the total number of actual positives

```{r auxx}
cens10 <- ifelse(y_test[,"time"]  < 10 & y_test[, "status"]== 0, 1, 0)

table(cens10)
ttx <- cbind(y_test[,"time"], y_test[, "status"], cens10)
colnames(ttx) <- c("time", 'status', "cens10")
ttx[1:10,]
Cp10_test <- sum(cens10)/length(cens10) 
nnc <- sum(cens10 ==0) # number of non-censored obs
nev <- sum(cens10 ==0 & y_test[, "status"]==1)
prev10_test <- nev/nnc  #prevalence
Cp_test <- Cp10_test
prev_test <- prev10_test

# Cleanup
Cp10_test <- prev10_test <- NULL
```

We extract model fit for a given $\alpha$ and store it in `cv_opt_model` object
for later use.

```{r cva-glmnet-fit}
names(res_cvfit)
cvfit <- res_cvfit$ALLx

pfit_aopt <- cvfit$glmnet
class(pfit_aopt)
names(pfit_aopt)
pfit_aopt$lambda

lmbda <- pfit_aopt$lambda
sel <- c(1, 15, 18, 24, 31) # 30 -> 31
nms <- paste0("step", sel)
lmbda_sel <-lmbda[sel]
names(lmbda_sel) <- nms
lmbda_sel
lmbda_opt <- lmbda_sel[1]
```

# Model Performance 




Predictive performance of a selected model using _standard_ approach.


## C-index

Step 1: Linear predictor estimation 

Estimate linear predictor values for the Cox model fitted to test data. 

```{r calc-pred2}
predM_lpmtx <- predict(pfit_aopt, newx = x_test, type = "link") # Matrix
```
 

Step2: C-index calculations

Ref: Harrel Jr, F. E. and Lee, K. L. and Mark, D. B. (1996) Tutorial in biostatistics:
multivariable prognostic models: issues in developing models, evaluating assumptions 
and adequacy, and measuring and reducing error, _Statistics in Medicine_, 15, pages 361-387.


Note: s0, ... s30 correspond to step1, ... step31

```{r C-index-Mod}
Cindex_lmbda <- apply(predM_lpmtx, 2, Cindex, y = ySurv_test) # Multiple lambda
length(Cindex_lmbda)

Cindex_lmbda[sel] 

#Cindex(predM_lp, ySurv_test)   # For optimal lambda
```

# Time-dependent ROC for various `step` values


* Initialize list with the results


```{r res-list}
lmbda_sel
len <- length(lmbda_sel)
survROC_list <- vector(mode ="list", length = len)
names(survROC_list) <- names(lmbda_sel)
str(survROC_list)
```

## Score (illustration)

Higher `BL_eGFR` -> lower score -> lower hazard of developing ESKD

```{r score-details}
tmp1 <- apply(x_test, 2, mean)
tmp2 <- tmp1
tmp2["BL_eGFR"] <- tmp1["BL_eGFR"]+10
x_toy <- rbind(tmp1,tmp2)
predM_lp <- as.vector(predict(pfit_aopt, newx = x_toy, type = "link", s = lmbda_opt))
res <-cbind(x_toy[, "BL_eGFR"], predM_lp) 
colnames(res) <- c("BL_eGFR", "lp_score")
res
```

<!--   ------  STEP1  ------ -->

```{r step1-roc, echo = FALSE}
step_sel <- "step1"
```

## ROC for `r step_sel`

* source: (https://datascienceplus.com/time-dependent-roc-for-survival-prediction-models-in-r)
* Youden W. J. Index for rating diagnostic tests. Cancer. 1950;3(1):32-35
* https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5470053/
* Pepe M. S. The Statistical Evaluation of Medical Tests for Classification and Prediction. Vol. 28. Oxford, UK: Oxford University Press; 2003. (Oxford Statistical Science Series)
* Perkins N. J., Schisterman E. F. The inconsistency of optimal cut-points using two ROC based criteria. American Journal of Epidemiology. 2006;163(7):670-675.
* Liu X. Classification accuracy and cut point selection. Statistics in Medicine. 2012;31(23):2676-2686. doi: 10.1002/sim.4509
* Ilkr Unal Defining an Optimal Cut-Point Value in ROC Analysis: An Alternative Approach Comput Math Methods Med. 2017; 2017: 3762651. Published online 2017 May 31.  

```{r surv-roc}
dataM_test <- test_data
dataM_test$predM_lp <- NULL


lmbda_opt <-  lmbda_sel[names(lmbda_sel) == step_sel]
predM_lp <- as.vector(predict(pfit_aopt, newx = x_test, type = "link", s = lmbda_opt))
# predM_lp <- - predM_lp # !!!
predM_lp[1:10]
summary(predM_lp)

                
## Augment `dataM_test` with linear predictor
 dataM_test$predM_lp <- predM_lp

# Evaluate every 2.5 years
 
tx <- 2.5* c(1,2,3,4) # ROC time points
tx
predM_lp[1:10]
survROC_lp <- create_survivalROC_data(tx, dataM_test, predM_lp)

survROC_lp2 <- survROC_lp %>% mutate(sens = TP, spec = 1- FP, 
               num = sens*prev_test, 
               den = sens*prev_test + (1-spec)*(1-prev_test),
               PPV = num/den,
               Youden = TP- FP,
               ER01 = sqrt(FP**2 +(1-TP)**2),  # The Closest to (0,1) Criteria (ER) Pepe (2003), Perkins (2006)
               Liu  =sens*spec,                # Concordance Probability Method (CZ), Liu 2012
               IU  = abs(sens -auc) + abs(spec -auc) # Index of Union (IU): Unal (2017)
         ) %>% select(-c(num, den))
 
survROC_lp2 %>% print (n=50)

```

Save results in a component of a list

```{r step1-res}
step_sel
survROC_list[[step_sel]]  <- survROC_lp2
survROC_lp2 <- NULL
```
* Plot Time-dependent ROC every 2.5 years

```{r surv-roc-plot}
## Plot Time-dependent ROC every 2.5 years
plot_timedep_ROC(survROC_lp) 
```

<!--   ------  STEP15  ------ -->

```{r step15-roc, echo = FALSE}
step_sel <- "step15"
```

## ROC for `r step_sel`

* source: (https://datascienceplus.com/time-dependent-roc-for-survival-prediction-models-in-r)


```{r surv-roc15}
dataM_test <- test_data
dataM_test$predM_lp <- NULL


lmbda_opt <-  lmbda_sel[names(lmbda_sel) == step_sel]
predM_lp <- as.vector(predict(pfit_aopt, newx = x_test, type = "link", s = lmbda_opt))
# predM_lp <- - predM_lp # !!!
predM_lp[1:10]
summary(predM_lp)

                
## Augment `dataM_test` with linear predictor
 dataM_test$predM_lp <- predM_lp

# Evaluate every 2.5 years
 
tx <- 2.5* c(1,2,3,4) # ROC time points
tx
predM_lp[1:10]
survROC_lp <- create_survivalROC_data(tx, dataM_test, predM_lp)

survROC_lp2 <- survROC_lp %>% mutate(sens = TP, spec = 1- FP, 
               num = sens*prev_test, 
               den = sens*prev_test + (1-spec)*(1-prev_test),
               PPV = num/den,
               Youden = TP- FP,
               ER01 = sqrt(FP**2 +(1-TP)**2),  # The Closest to (0,1) Criteria (ER) Pepe (2003), Perkins (2006)
               Liu  =sens*spec,                # Concordance Probability Method (CZ), Liu 2012
               IU  = abs(sens -auc) + abs(spec -auc) # Index of Union (IU): Unal (2017)
               
         ) %>% select(-c(num, den))
 
survROC_lp2 %>% print (n=50)

```

Save results in a component of a list

```{r step15-res}
step_sel
survROC_list[[step_sel]]  <- survROC_lp2
survROC_lp2 <- NULL
```
* Plot Time-dependent ROC every 2.5 years

```{r surv-roc15-plot}
## Plot Time-dependent ROC every 2.5 years
plot_timedep_ROC(survROC_lp) 
```
<!--   ------  STEP18  ------ -->

```{r step18-roc, echo = FALSE}
step_sel <- "step18"
```

## ROC for `r step_sel`

* source: (https://datascienceplus.com/time-dependent-roc-for-survival-prediction-models-in-r)


```{r surv-roc18}
dataM_test <- test_data
dataM_test$predM_lp <- NULL


lmbda_opt <-  lmbda_sel[names(lmbda_sel) == step_sel]
predM_lp <- as.vector(predict(pfit_aopt, newx = x_test, type = "link", s = lmbda_opt))
# predM_lp <- - predM_lp # !!!
predM_lp[1:10]
summary(predM_lp)

                
## Augment `dataM_test` with linear predictor
 dataM_test$predM_lp <- predM_lp

# Evaluate every 2.5 years
 
tx <- 2.5* c(1,2,3,4) # ROC time points
tx
predM_lp[1:10]
survROC_lp <- create_survivalROC_data(tx, dataM_test, predM_lp)

survROC_lp2 <- survROC_lp %>% mutate(sens = TP, spec = 1- FP, 
               num = sens*prev_test, 
               den = sens*prev_test + (1-spec)*(1-prev_test),
               PPV = num/den,
               Youden = TP- FP,
               ER01 = sqrt(FP**2 +(1-TP)**2),  # The Closest to (0,1) Criteria (ER) Pepe (2003), Perkins (2006)
               Liu  =sens*spec,                # Concordance Probability Method (CZ), Liu 2012
               IU  = abs(sens -auc) + abs(spec -auc) # Index of Union (IU): Unal (2017)
         ) %>% select(-c(num, den))
 
survROC_lp2 %>% print (n=50)

```

Save results in a component of a list

```{r step18-res}
step_sel
survROC_list[[step_sel]]  <- survROC_lp2
survROC_lp2 <- NULL
```
* Plot Time-dependent ROC every 2.5 years

```{r surv-roc18-plot}
## Plot Time-dependent ROC every 2.5 years
plot_timedep_ROC(survROC_lp) 
```

<!--   ------  STEP24  ------ -->

```{r step24-roc, echo = FALSE}
step_sel <- "step24"
```

## ROC for `r step_sel`

* source: (https://datascienceplus.com/time-dependent-roc-for-survival-prediction-models-in-r)


```{r surv-roc24}
dataM_test <- test_data
dataM_test$predM_lp <- NULL

lmbda_opt <-  lmbda_sel[names(lmbda_sel) == step_sel]
predM_lp <- as.vector(predict(pfit_aopt, newx = x_test, type = "link", s = lmbda_opt))
# predM_lp <- - predM_lp # !!!
predM_lp[1:10]
summary(predM_lp)

                
## Augment `dataM_test` with linear predictor
 dataM_test$predM_lp <- predM_lp

# Evaluate every 2.5 years
 
tx <- 2.5* c(1,2,3,4) # ROC time points
tx
predM_lp[1:10]
survROC_lp <- create_survivalROC_data(tx, dataM_test, predM_lp)

survROC_lp2 <- survROC_lp %>% mutate(sens = TP, spec = 1- FP, 
               num = sens*prev_test, 
               den = sens*prev_test + (1-spec)*(1-prev_test),
               PPV = num/den,
               Youden = TP- FP,
               ER01 = sqrt(FP**2 +(1-TP)**2),  # The Closest to (0,1) Criteria (ER) Pepe (2003), Perkins (2006)
               Liu  =sens*spec,                # Concordance Probability Method (CZ), Liu 2012
               IU  = abs(sens -auc) + abs(spec -auc) # Index of Union (IU): Unal (2017)

         ) %>% select(-c(num, den))
 
survROC_lp2 %>% print (n=50)

```

Save results in a component of a list

```{r step24-res}
step_sel
survROC_list[[step_sel]]  <- survROC_lp2
survROC_lp2 <- NULL
```
* Plot Time-dependent ROC every 2.5 years

```{r surv-roc24-plot}
## Plot Time-dependent ROC every 2.5 years
plot_timedep_ROC(survROC_lp) 
```



<!--   ------  STEP31  ------ -->

```{r step31-roc, echo = FALSE}
step_sel <- "step31"
```

## ROC for `r step_sel`


* source: (https://datascienceplus.com/time-dependent-roc-for-survival-prediction-models-in-r)


```{r surv-roc31}
dataM_test <- test_data
dataM_test$predM_lp <- NULL
lmbda_opt <-  lmbda_sel[names(lmbda_sel) == step_sel]
predM_lp <- as.vector(predict(pfit_aopt, newx = x_test, type = "link", s = lmbda_opt))
# predM_lp <- - predM_lp # !!!
predM_lp[1:10]
summary(predM_lp)

                
## Augment `dataM_test` with linear predictor
 dataM_test$predM_lp <- predM_lp

# Evaluate every 2.5 years
 
tx <- 2.5* c(1,2,3,4) # ROC time points
tx
predM_lp[1:10]
survROC_lp <- create_survivalROC_data(tx, dataM_test, predM_lp)
survROC_lp2 <- survROC_lp %>% mutate(sens = TP, spec = 1- FP, 
               num = sens*prev_test, 
               den = sens*prev_test + (1-spec)*(1-prev_test),
               PPV = num/den,
               Youden = TP- FP,
               ER01 = sqrt(FP**2 +(1-TP)**2),  # The Closest to (0,1) Criteria (ER) Pepe (2003), Perkins (2006)
               Liu  =sens*spec,                # Concordance Probability Method (CZ), Liu 2012
               IU  = abs(sens -auc) + abs(spec -auc) # Index of Union (IU): Unal (2017)

         ) %>% select(-c(num, den))
 
survROC_lp2 %>% print (n=50)

```

Save results in a component of a list

```{r step31-res}
step_sel
survROC_list[[step_sel]]  <- survROC_lp2
survROC_lp2 <- NULL
```



* Plot Time-dependent ROC every 2.5 years

```{r surv-roc-plot31}
## Plot Time-dependent ROC every 2.5 years
plot_timedep_ROC(survROC_lp) 
```

Save results in Excel file

```{save-xlsx}
library(writexl)
str(survROC_list)
xlsxf <- "survROC_list.xlsx"
xlsxp <- paste0("./validation_save/", xlsxf)
write_xlsx(survROC_list, xlsxp)

```



```{r exit0, include=FALSE}
knitr::knit_exit()
```
