---
title: '6.Cox Univariate Analysis'
author: "Chunyi Wu and Andrzej Galecki"
date: "`r as.character(Sys.Date(), format = '%A %B %d, %Y')`"
output:
  rmdformats::readthedown:
     lightbox: true
     use_bookdown: true
---


```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```


```{r pacman, echo=FALSE, message=FALSE, warning=FALSE}
if (!require('pacman')) install.packages('pacman', repos = "http://cran.us.r-project.org")
library(pacman)

pacman::p_load(
  readxl,       # load excel file
  glmnet,       # Lasso and Elastic-Net Regularized Generalized Linear Models
  rmdformats,   # rmd formats
  rmarkdown,    # rmarkdown
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics 
  janitor,      # adding totals and percents to tables
  flextable,    # format the output table
  gtsummary,    # calculate summary statistics and format the results
  sjPlot,       # correlation matrix
  purrr,        # enhances R functional programming (FP) toolkit 
  tidyr,        #Tools to help to create tidy data
  ggplot2,      #Plot the results
  corrplot,     #correlation plot
  survival      # survival analysis    
  )
```

# R setup

```{r  rsetup}
sessionInfo()
```


```{r, data-read, echo=FALSE, message=FALSE, warning=FALSE}
xlsx_path = "./Data/data_example.xlsx"

scrambled_orig=readxl::read_excel(xlsx_path, na = "", guess_max = 1000)
scrambled <- scrambled_orig  %>% mutate(log10_B_HBA1C_PRC=log10(B_HBA1C_PRC),log10_DU_ACR=log10(DU_ACR)) %>%
    filter(BL_eGFR >=45)%>%
mutate(CASE_CONTROL = if_else(FU_TIME >= 10, 0, CASE_CONTROL),
       FU_TIME = if_else(FU_TIME > 10, 10, FU_TIME))
       #administrative censoring at 10 years
```
Raw data stored in `r xlsx_path` file contain `r nrow(scrambled_orig)` rows.  
Variable names are listed below.

```{r var-names}
names(scrambled_orig)
```


```{r, table1, echo=FALSE, message=FALSE, warning=FALSE}
scrambled2 <- scrambled  %>% 
          mutate(Gender=case_when(SEX==1 ~"Male",SEX==2 ~"Female"),
              Status=case_when(CASE_CONTROL==1 ~"Case",CASE_CONTROL==0 ~"Control"))%>%                    
          select( "Status","FU_TIME","CASE_CONTROL","BL_eGFR","B_HBA1C_PRC","log10_DU_ACR","Gender","AGE_TL","TNF.R1.npx","TNF.R2.npx","LTBR.npx","TNFRSF4.npx","TNFRSF6B.npx","CD27.npx","TNFRSF10A.npx","RELT.npx","EDA2R.npx","WFDC2.npx", "PI3_npx","CD160.npx","EPHA2.npx","EFNA4.npx","GFR_alpha_1_npx","DLL1.npx","LAYN.npx","SYND1.npx","KIM1.npx","PVRL4.npx","IL.1RT1.npx" )      

scrambled2$FU_TIME=as.numeric(scrambled2$FU_TIME)
```

# Administrative censoring at 10 years
```{r, table2, echo=FALSE, message=FALSE, warning=FALSE}

scrambled2_ADMIN <- scrambled2  %>% select(Status,FU_TIME)

scrambled2_ADMIN %>%
  tbl_summary(by = Status, missing_text = "(Missing N)",   label = list(FU_TIME ~ "Follow-up Time (year)"), type = all_continuous() ~ "continuous2", statistic = list(all_continuous() ~ c("{mean}, {sd}","{median} ({p25}, {p75})", "{min}, {max}"),all_categorical() ~ "{n} ({p}%)"))  %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>%
  add_overall() %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Status**") %>%
  modify_footnote( all_stat_cols() ~ "Mean (SD), Median (IQR) or Frequency (%)")%>%
  modify_caption("**Follow-up time by using administrative censoring at 10 years**") %>%
  bold_labels()
```

# Univariate statistics - Cox Model
```{r, table3, echo=FALSE, message=FALSE, warning=FALSE}

cox_univariate <-
  tbl_uvregression(
    scrambled2[c("FU_TIME", "CASE_CONTROL", "BL_eGFR","B_HBA1C_PRC","log10_DU_ACR","Gender","AGE_TL","TNF.R1.npx","TNF.R2.npx","LTBR.npx","TNFRSF4.npx","TNFRSF6B.npx","CD27.npx","TNFRSF10A.npx","RELT.npx","EDA2R.npx","WFDC2.npx", "PI3_npx","CD160.npx","EPHA2.npx","EFNA4.npx","GFR_alpha_1_npx","DLL1.npx","LAYN.npx","SYND1.npx","KIM1.npx","PVRL4.npx","IL.1RT1.npx")],
    method = coxph,
    y = Surv(FU_TIME, CASE_CONTROL),
    exponentiate = TRUE,
    pvalue_fun = function(x) style_pvalue(x, digits = 2)
  )

cox_univariate


```

# C-index
```{r, table4, echo=FALSE, message=FALSE, warning=FALSE}

calculate_concordance <- function(predictor, data) {
  # Build the Cox model formula
  formula <- as.formula(paste("Surv(FU_TIME, CASE_CONTROL) ~", predictor))
  
  # Fit the Cox model
  fit <- coxph(formula, data = data)
  
  # Calculate the concordance
  res <- concordance(fit, timewt = "n")
  
  return(res$concordance)
}

# Predictor variables
predictors <- c("BL_eGFR","B_HBA1C_PRC","log10_DU_ACR","Gender","AGE_TL","TNF.R1.npx","TNF.R2.npx","LTBR.npx","TNFRSF4.npx","TNFRSF6B.npx","CD27.npx","TNFRSF10A.npx","RELT.npx","EDA2R.npx","WFDC2.npx", "PI3_npx","CD160.npx","EPHA2.npx","EFNA4.npx","GFR_alpha_1_npx","DLL1.npx","LAYN.npx","SYND1.npx","KIM1.npx","PVRL4.npx","IL.1RT1.npx")


# Calculate concordance for each predictor
results2 <- sapply(predictors, calculate_concordance, data = scrambled2)

# Print the concordance results
results_df <- data.frame(C_index = results2)

# Print the results data frame
print(results_df)