---
title: '7.Logistic Univariate Analysis'
author: "Chunyi Wu and Andrzej Galecki"
date: "`r as.character(Sys.Date(), format = '%A %B %d, %Y')`"
output:
  rmdformats::readthedown:
     lightbox: true
     use_bookdown: true
---


```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```


```{r pacman, echo=FALSE, message=FALSE, warning=FALSE}
if (!require('pacman')) install.packages('pacman', repos = "http://cran.us.r-project.org")
library(pacman)

pacman::p_load(
  readxl,       # load excel file
  glmnet,       # Lasso and Elastic-Net Regularized Generalized Linear Models
  rmdformats,   # rmd formats
  rmarkdown,    # rmarkdown
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics 
  janitor,      # adding totals and percents to tables
  flextable,    # format the output table
  gtsummary,    # calculate summary statistics and format the results
  sjPlot,       # correlation matrix
  purrr,        # enhances R functional programming (FP) toolkit 
  tidyr,        # Tools to help to create tidy data
  ggplot2,      # Plot the results
  corrplot,     # correlation plot
  survival,     # survival analysis,    
  pROC          # ROC for c-statistics
  )
```

# R setup

```{r  rsetup}
sessionInfo()
```


```{r, data-read, echo=FALSE, message=FALSE, warning=FALSE}
xlsx_path = "./Data/data_example.xlsx"

scrambled_orig=readxl::read_excel(xlsx_path, na = "", guess_max = 1000)
scrambled <- scrambled_orig  %>% mutate(log10_DU_ACR=log10(DU_ACR),eGFR_by_10_mL=BL_eGFR/10) %>%
    filter(BL_eGFR >=45)%>%
    filter(FU_TIME > 0)
# mutate(CASE_CONTROL = if_else(FU_TIME >= 10, 0, CASE_CONTROL),
#        FU_TIME = if_else(FU_TIME >= 10, 10, FU_TIME))
       #administrative censoring at 10 years
       # %>%filter(CASE_CONTROL==1 | (CASE_CONTROL==0 & FU_TIME>=5))
```
Raw data stored in `r xlsx_path` file contain `r nrow(scrambled_orig)` rows.  
Variable names are listed below.

```{r var-names}
names(scrambled_orig)
```


```{r, table1, echo=FALSE, message=FALSE, warning=FALSE}
scrambled2 <- scrambled  %>% 
          mutate(Gender=case_when(SEX==1 ~"Male",SEX==2 ~"Female"),
              Status=case_when(CASE_CONTROL==1 ~"Case",CASE_CONTROL==0 ~"Control"))%>%                    
          select( "Status","FU_TIME","CASE_CONTROL","BL_eGFR","eGFR_by_10_mL","B_HBA1C_PRC","log10_DU_ACR","DU_ACR","Gender","AGE_TL","TNF.R1.npx","TNF.R2.npx","LTBR.npx","TNFRSF4.npx","TNFRSF6B.npx","CD27.npx","TNFRSF10A.npx","RELT.npx","EDA2R.npx","WFDC2.npx", "PI3_npx","CD160.npx","EPHA2.npx","EFNA4.npx","GFR_alpha_1_npx","DLL1.npx","LAYN.npx","SYND1.npx","KIM1.npx","PVRL4.npx","IL.1RT1.npx" )      

scrambled2$FU_TIME=as.numeric(scrambled2$FU_TIME)
```


# Follo-up time for a case-control study
```{r, table2, echo=FALSE, message=FALSE, warning=FALSE}

scrambled2_ADMIN <- scrambled2  %>% select(Status,FU_TIME)

scrambled2_ADMIN %>%
  tbl_summary(by = Status, missing_text = "(Missing N)",   label = list(FU_TIME ~ "Follow-up Time (year)"), type = all_continuous() ~ "continuous2", statistic = list(all_continuous() ~ c("{mean}, {sd}","{median} ({p25}, {p75})", "{min}, {max}"),all_categorical() ~ "{n} ({p}%)"))  %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>%
  add_overall() %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Status**") %>%
  modify_footnote( all_stat_cols() ~ "Mean (SD), Median (IQR) or Frequency (%)")%>%
  modify_caption("**Follow-up time by using administrative censoring at 10 years**") %>%
  bold_labels()
```

# Univariate statistics - Logistic Model
```{r, table3, echo=FALSE, message=FALSE, warning=FALSE}

Logistic_univariate <-
  tbl_uvregression(
        scrambled2[c("CASE_CONTROL", "eGFR_by_10_mL","B_HBA1C_PRC","log10_DU_ACR","Gender","AGE_TL","TNF.R1.npx","TNF.R2.npx","LTBR.npx","TNFRSF4.npx","TNFRSF6B.npx","CD27.npx","TNFRSF10A.npx","RELT.npx","EDA2R.npx","WFDC2.npx", "PI3_npx","CD160.npx","EPHA2.npx","EFNA4.npx","GFR_alpha_1_npx","DLL1.npx","LAYN.npx","SYND1.npx","KIM1.npx","PVRL4.npx","IL.1RT1.npx")],
    method = glm,
    y = CASE_CONTROL,
    method.args = list(family = binomial),
    exponentiate = TRUE
  )

Logistic_univariate


```

# C-index
```{r, table4, echo=FALSE, message=FALSE, warning=FALSE}

calculate_concordance <- function(predictor, data) {
  # Build the logistic regression model formula
  formula <- as.formula(paste("CASE_CONTROL ~", predictor))
  
  # Fit the logistic regression model
  fit <- glm(formula, data = data, family = binomial)
  
  # Calculate the concordance
  res <- concordance(fit)
  
  # Create a data frame with the predictor name, C-index, and C-index SE
  result_df <- data.frame(Predictor = predictor,
                          C_index = res$concordance,
                          C_index_se = sqrt(res$var))
  
  return(result_df)
}

# Predictor variables
predictors_var <- c("eGFR_by_10_mL", "B_HBA1C_PRC", "log10_DU_ACR", "Gender", "AGE_TL", "TNF.R1.npx", "TNF.R2.npx", "LTBR.npx", "TNFRSF4.npx", "TNFRSF6B.npx", "CD27.npx", "TNFRSF10A.npx", "RELT.npx", "EDA2R.npx", "WFDC2.npx", "PI3_npx", "CD160.npx", "EPHA2.npx", "EFNA4.npx", "GFR_alpha_1_npx", "DLL1.npx", "LAYN.npx", "SYND1.npx", "KIM1.npx", "PVRL4.npx", "IL.1RT1.npx")

# Calculate concordance for each predictor
results <- lapply(predictors_var, calculate_concordance, data = scrambled2)

# Combine the results into a single data frame
results_df <- do.call(rbind, results)

# Print the results data frame
print(results_df)

```

# Univariate statistics by Categorical Quartile- Logistic Model
```{r, table5, echo=FALSE, message=FALSE, warning=FALSE}

# Create a list of predictors
predictors <- c("BL_eGFR","B_HBA1C_PRC","DU_ACR","AGE_TL","TNF.R1.npx","TNF.R2.npx","LTBR.npx","TNFRSF4.npx","TNFRSF6B.npx","CD27.npx","TNFRSF10A.npx","RELT.npx","EDA2R.npx","WFDC2.npx", "PI3_npx","CD160.npx","EPHA2.npx","EFNA4.npx","GFR_alpha_1_npx","DLL1.npx","LAYN.npx","SYND1.npx","KIM1.npx","PVRL4.npx","IL.1RT1.npx")

# Loop through each predictor
for (predictor in predictors) {
  # Create quartile variable name
  quartile_var <- paste0(predictor, "_quartile")
  
  # Calculate quartiles
  quartiles <- quantile(scrambled2[[predictor]], probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
  
  # Assign quartile values to the quartile variable
  scrambled2[[quartile_var]] <- cut(scrambled2[[predictor]], breaks = quartiles, include.lowest = TRUE, labels = c(1, 2, 3, 4))
}

Logistic_univariate <-
  tbl_uvregression(
    scrambled2[c("CASE_CONTROL", "BL_eGFR_quartile","B_HBA1C_PRC_quartile","DU_ACR_quartile","AGE_TL_quartile","TNF.R1.npx_quartile","TNF.R2.npx_quartile","LTBR.npx_quartile","TNFRSF4.npx_quartile","TNFRSF6B.npx_quartile","CD27.npx_quartile","TNFRSF10A.npx_quartile","RELT.npx_quartile","EDA2R.npx_quartile","WFDC2.npx_quartile", "PI3_npx_quartile","CD160.npx_quartile","EPHA2.npx_quartile","EFNA4.npx_quartile","GFR_alpha_1_npx_quartile","DLL1.npx_quartile","LAYN.npx_quartile","SYND1.npx_quartile","KIM1.npx_quartile","PVRL4.npx_quartile","IL.1RT1.npx_quartile")],
    method = glm,
    y = CASE_CONTROL,
    method.args = list(family = binomial),
    exponentiate = TRUE
  )

Logistic_univariate

```


# C-index by Categorical Quartile
```{r, table6, echo=FALSE, message=FALSE, warning=FALSE}

# Predictor variables
predictors <- c("BL_eGFR_quartile", "B_HBA1C_PRC_quartile", "DU_ACR_quartile", "AGE_TL_quartile","TNF.R1.npx_quartile", "TNF.R2.npx_quartile", "LTBR.npx_quartile","TNFRSF4.npx_quartile","TNFRSF6B.npx_quartile", "CD27.npx_quartile", "TNFRSF10A.npx_quartile","RELT.npx_quartile","EDA2R.npx_quartile", "WFDC2.npx_quartile", "PI3_npx_quartile", "CD160.npx_quartile","EPHA2.npx_quartile", "EFNA4.npx_quartile", "GFR_alpha_1_npx_quartile", "DLL1.npx_quartile","LAYN.npx_quartile", "SYND1.npx_quartile", "KIM1.npx_quartile", "PVRL4.npx_quartile","IL.1RT1.npx_quartile")

# Calculate concordance for each predictor
results2 <- lapply(predictors, calculate_concordance, data = scrambled2)

# Combine the results into a single data frame
results_df2 <- do.call(rbind, results2)

# Print the results data frame
print(results_df2)
```


# Univariate statistics by Numerical Quartile- Logistic Model
```{r, table7, echo=FALSE, message=FALSE, warning=FALSE}

#recode factor variables to numerical variables
recoded_vars <- c("BL_eGFR_quartile","B_HBA1C_PRC_quartile","DU_ACR_quartile","AGE_TL_quartile","TNF.R1.npx_quartile","TNF.R2.npx_quartile","LTBR.npx_quartile","TNFRSF4.npx_quartile","TNFRSF6B.npx_quartile","CD27.npx_quartile","TNFRSF10A.npx_quartile","RELT.npx_quartile","EDA2R.npx_quartile","WFDC2.npx_quartile", "PI3_npx_quartile","CD160.npx_quartile","EPHA2.npx_quartile","EFNA4.npx_quartile","GFR_alpha_1_npx_quartile","DLL1.npx_quartile","LAYN.npx_quartile","SYND1.npx_quartile","KIM1.npx_quartile","PVRL4.npx_quartile","IL.1RT1.npx_quartile")

for (var in recoded_vars) {
  scrambled2[[var]] <- as.numeric(as.character(scrambled2[[var]]))
}

Logistic_univariate <-
  tbl_uvregression(
    scrambled2[c("CASE_CONTROL", "BL_eGFR_quartile","B_HBA1C_PRC_quartile","DU_ACR_quartile","AGE_TL_quartile","TNF.R1.npx_quartile","TNF.R2.npx_quartile","LTBR.npx_quartile","TNFRSF4.npx_quartile","TNFRSF6B.npx_quartile","CD27.npx_quartile","TNFRSF10A.npx_quartile","RELT.npx_quartile","EDA2R.npx_quartile","WFDC2.npx_quartile", "PI3_npx_quartile","CD160.npx_quartile","EPHA2.npx_quartile","EFNA4.npx_quartile","GFR_alpha_1_npx_quartile","DLL1.npx_quartile","LAYN.npx_quartile","SYND1.npx_quartile","KIM1.npx_quartile","PVRL4.npx_quartile","IL.1RT1.npx_quartile")],
    method = glm,
    y = CASE_CONTROL,
    method.args = list(family = binomial),
    exponentiate = TRUE
  )

Logistic_univariate

```


# C-index by Numerical Quartile
```{r, table8, echo=FALSE, message=FALSE, warning=FALSE}

# Predictor variables
predictors <- c("BL_eGFR_quartile", "B_HBA1C_PRC_quartile", "DU_ACR_quartile", "AGE_TL_quartile",
                "TNF.R1.npx_quartile", "TNF.R2.npx_quartile", "LTBR.npx_quartile", "TNFRSF4.npx_quartile","TNFRSF6B.npx_quartile", "CD27.npx_quartile", "TNFRSF10A.npx_quartile", "RELT.npx_quartile","EDA2R.npx_quartile", "WFDC2.npx_quartile", "PI3_npx_quartile", "CD160.npx_quartile",
"EPHA2.npx_quartile", "EFNA4.npx_quartile", "GFR_alpha_1_npx_quartile", "DLL1.npx_quartile","LAYN.npx_quartile", "SYND1.npx_quartile", "KIM1.npx_quartile", "PVRL4.npx_quartile",
 "IL.1RT1.npx_quartile")

# Calculate concordance for each predictor
results3 <- lapply(predictors, calculate_concordance, data = scrambled2)

# Combine the results into a single data frame
results_df3 <- do.call(rbind, results3)

# Print the results data frame
print(results_df3)
```